---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Kyle Knaut"
date: "January 19th, 2024"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r characterizing_Dennys}
#The Denny's dataset has 1643 observations (AKA 1643 Denny's)
print(nrow(dennys))

#There are 6 variables in this dataset
print(ncol(dennys))

#The variables in this dataset are shown below
names(dennys)

```
### Exercise 2
```{r characterizing_La_Quintas}
#The La Quinta dataset has 909 observations (AKA 909 La Quintas)
print(nrow(laquinta))

#There are 6 variables in this dataset
print(ncol(laquinta))

#The variables in this dataset are shown below
names(laquinta)

```

### Exercise 3

There are no Dennys locations outside of the US according to their website. There are, however, many La Quintas outside of the United States, with locations in Canada, Mexico, China, New Zealand, Turkey, The United Arab Emirates, Chile, Colombia, and Ecuador

### Exercise 4

In order to determine where locations outside of the us are within the data, I would use the filter function. I initially thought would filter the dataset for values where country != "US", however, after looking, there is no country variable for these datasets. Instead, I will look into the data and see how locations outside of the US code their zipcodes, or states. If either of these show up as NA for countries other than the US, then I will filter out these null values, hopefully resulting in an analysis of domestic locations. 

Okay, so after trolling through the data trying to find a La Quinta in another country, I have realized that this may be more difficult than I had initially thought. I have found two La Quintas in Puebla, Mexico, but both of them have a value for state (PU), and a zipcode. The only other ways I could think of to do this are by filtering for the 51 states in the US and dropping all other values, something with longitude and lattitude (but this is complicated with Alaska and Hawaii), or by using the states dataset somehow, maybe using that set to filter for the 51 states. 

### Exercise 5
It's the function I was dreaming of! Still no Dennys outside the US but cool the function exists :)
```{r Dennys_by_country}
dennys %>%
  filter(!(state %in% states$abbreviation))
```

### Exercise 6

``` {r Dennys_country_variable}
dennys %>%
  mutate(country = "US")
```

### Exercise 7
``` {r laquinta_countries_and_states}

#Canada (ON, BC)
#Colombia (ANT)
#Mexico (AG, CH, PU, VE, SL)
#No Observation for Quito, Ecuador
#No Observation for Santiago, Chile
#No Observation for Dubai, UAE
#No Observations for Bodrum, Giresun, or Istanbul, Turkey
#No Observations for Auckland or Queenstown, New Zealand
#No Observations for Weifang or Taiyuan, China
```

### Exercise 8
```{r Adding_country_variable}
laquinta %>%
  mutate(country = case_when(state %in% state.abb ~ "United States", state %in% c("ON", "BC") ~ "Canada", state == "ANT" ~ "Colombia", state %in% c("AG", "CH", "PU", "VE", "SL") ~ "Mexico"))

#I feel like we could use != to do this with less headache but still 
#glad for the experience

#laquinta <- laquinta %>%
  #filter(country == "United States")
#Did not recognize country as an object despite the fact that it showed
#up perfectly in my table

#exists(country)
laquinta <- laquinta %>%
  filter(state %in% states$abbreviation)
```

### Exercise 9
```{r count_states}
laquinta %>%
  count(state, sort = TRUE)

dennys %>%
  count(state, sort = TRUE)
```
Dennys:
  California has the most Denny's by far, with 203 more than the next closest state, which is Texas (who have 200 Denny's).
  Delaware has the least Denny's of any state with only one Dennys. Poor Delaware, Grand Slams are rare there :((
  
La Quinta:
  Texas comes in at number one here, with 237 La Quinta Inns. This is far more than any other state, as the next highest state is Florida with 74.
  Maine has the fewest La Quinta Inns, with only one :(
  
All in all, the results of this cursory glance make a lot of sense, but they do, unfortunately, violate Mitch Hedburg's premise that La Quinta roughly translates to next to the Dennys, especially in California, where there are nearly 8 times as many Denny's as there are La Quinta Inns.

### Exercise 10
```{r merging_datasets}
dennys <- dennys %>%
  mutate(establishment = "Denny's")

laquinta <- laquinta %>%
  mutate(establishment = "La Quinta")

dennys_laquinta <- bind_rows(dennys, laquinta)

ggplot(dennys_laquinta, mapping = aes(x = longitude, y = latitude, color = establishment)) + geom_point() + labs(x = "Longitude", y = "Latitude", color = "Establishment", title = "Distribution of Denny's and La Quinta Inns", subtitle = "Within the United States")

```

### Exercise 11
```{r NC_Dennys_vs_LaQuinta}
dennys_laquinta %>%
  filter(state == "NC") %>%
  ggplot(dennys_laquinta, mapping = aes(x = longitude, y = latitude, color = establishment, alpha = 0.5)) + geom_point() + labs(x = "Longitude", y = "Latitude", color = "Establishment", title = "Distribution of Denny's and La Quinta Inns", subtitle = "Within North Carolina")
```

### Exercise 12
```{r Texas_Dennys_vs_LaQuinta}
dennys_laquinta %>%
  filter(state == "TX") %>%
  ggplot(dennys_laquinta, mapping = aes(x = longitude, y = latitude, color = establishment, alpha = 0.5)) + geom_point() + labs(x = "Longitude", y = "Latitude", color = "Establishment", title = "Distribution of Denny's and La Quinta Inns", subtitle = "Within Texas")

```
Mitch Hedberg's joke appears to hold a lot better here,  but it's hard to say for certain without getting into the nitty gritty stats.